---
title: "Script using Superstore sample data"
author: "Monika Aulich-Olszewska"
date: "08-08-2024"
---

This is an example script to show my R skills.
Used data is from Kaggler website.

On the first part of the script used libraries are written. To install libraries, use command install.packages("library_name")
```{r}
library(ggplot2)
library(VGAM) 

```

Next step is to read the data. It is important to choose right path to the file.

```{r}
csv_file <- read.csv("/Users/monia/Desktop/R/training/Sample-Superstore.csv", header = TRUE)
df <- data.frame(csv_file)

#to check if data is loaded one can use head() function
head(df)
df
```

Next step is to choose interesting data and what to do with them.
At first I want to seew what product is the most bought

```{r}
ggplot(df, aes(x=Product.ID, y=Row.ID)) + geom_bar(stat = "identity", color="lightpink") + labs(x="Product ID",y="Number of products", title="Bar plot with numbers of sold products")


```
As we can see, this plot won't tell us much. Let's try wider cathegory - Segment.

```{r}
ggplot(df, aes(x=Segment, y=Row.ID)) + geom_bar(stat = "identity", color="lightblue") + labs(x="Cathegory name",y="Number of sold products in segment", title="Bar plot with numbers of sold products")

```
So right now one can see, that the biggest group in Superstore is Consumer group. Let's see what info we can get from Home Office data.

First step here is to get data from Home office group only.

```{r}
df_ho <- df[df$Segment=="Home Office",]
head(df_ho)
print(paste("Number of sold products for Home Office Group is",nrow(df_ho)))
```

Now group is much smaller, so it is easier to find interesting data.
```{r}
ggplot(df_ho, aes(x=Ship.Mode, y=Row.ID)) + geom_bar(stat = "identity", color="lightpink") + labs(x="Product ID",y="Number of products", title="Bar plot with chosen ship mode")

```
As we can see the most used ship mode is Standard Class.
I would like to see if there is a correlation between used ship mode and product category. 
I used easy function cor()
```{r}
df_corr <- data.frame(df_ho$Ship.Mode,df_ho$Category)
colnames(df_corr) <- c("Ship.Mode","Category")

#only numeric variables can be used to computate correlation
df_corr$Ship_num <- as.numeric(as.factor(df_corr$Ship.Mode))
df_corr$Categ_num <- as.numeric(as.factor(df_corr$Category))

cor_ship_cat <- cor(df_corr$Ship_num,df_corr$Categ_num)

print(paste("Correlation is", round(cor_ship_cat,3)))

```
The correlation is very close to 0, so one can say that there is no correlation between used ship mode and product category.


What interesting is for me, I would also like to see, how many orders were in what year in chosen group.
At first I need to change data format
```{r}
df_ho$newdate <- strptime(as.character(df_ho$Order.Date), "%m/%d/%Y")
df_ho$Date <- format(df_ho$newdate, "%Y-%m-%d")
                          
splitted_data <- split(df_ho, format(as.Date(df_ho$Date,format="%Y-%m-%d"), format = "%Y"))

df_dates <- data.frame()
for(i in 1:length(splitted_data)){
  data <- splitted_data[[i]]
  year <- format(as.Date(data$Date[1], format="%Y-%m-%d"),"%Y")
  print(paste("In year",year[1],"there was",nrow(data),"solds"))
  line <- c(year,nrow(data))
  df_dates <- rbind(df_dates,line)
}
colnames(df_dates) <- c("Year","Solds")
#df_dates$Year <- sapply(df_dates$Year,as.numeric)

ggplot(data=df_dates, aes(x=Year, y=Solds, group=1)) + geom_line(color="pink")+ geom_point() + labs(x="Years",y="Number of sold products", title="Plot with number of sold products per year")

```
As we can see, every year there is more sold products in Home Office category.
